---
title: "color patch trimmed"
author: "Andrew O. Rubio"
date: "9/10/2020"
output: html_document
---

```{r}
library(tximport)
library(DESeq2)
library(dplyr)
library(foreach)
library(data.table)
library(splines)
library(ggthemes)
library(scales)
library(gridExtra)
library(tidyr)
library(pheatmap)
library(RColorBrewer)
library(ggplot2)
library("BiocParallel")
register(SnowParam(8))
library(apeglm)
library(tidyverse)
library(topGO)
library(GO.db)
library(tidyverse)
library(SummarizedExperiment)
library(pcaExplorer)
```


Some functions for this analysis.

```{r}
# this function pulls out particular a priori genes of interest into a new data frame
aprioris <- function(df, annos){
  df.aprioris <- data.frame()
  for (i in 1:length(colors$gene_name)){
    searchterm <- paste0("\\b", colors$gene_name[i], "\\b")
    searchterm <- tolower(searchterm)
    tmp <- df %>% filter(str_detect(Gene, searchterm))
    df.aprioris <- rbind(df.aprioris, tmp)
}
   return(df.aprioris)
}

# this function extracts all genes below a specified alpha value, prints the number of DE genes to screen, and saves a spreadsheet to file
siggies <- function(df, alpha, csv){
  sigs <- df %>% filter(padj < alpha)
  print(paste0("Number of significant genes: ", length(sigs$padj)))
  write.csv(sigs, csv, row.names = FALSE)
  return(sigs)
}


# significant gene wrapper function
SigGeneWrapper <- function(model_output, alpha, comparison){
  print(paste0("Running ", comparison, " with alpha = ", alpha))
  df <- setDT(as.data.frame(model_output), keep.rownames = "Gene")
  # add annotation data
  #df1 <- dplyr::left_join(df, annos, by = "transcript") 
  # get significant genes
  print("Overall significant genes")
  sigs <- siggies(df, alpha, paste0("results/results/", comparison, "_genes.csv"))
  # add annotation data
  colordf <- aprioris(df, annos)
  #get significant color genes
  print("Significant color genes")
  color.sigs <- siggies(colordf, alpha, paste0("results/results/", comparison, "_colorgenes.csv"))
}

```

#Data import

```{r}
sessionInfo()
```

#Data import

create sample spreadsheet.

```{r}
###### list all samples from expression data ####
# get the directory/path for each sample in this study
base_dir <- getwd()

filenames <- list.files(path = "gene counts/", full.names = F, recursive = F)

files <- file.path(base_dir, "gene counts/", filenames) # files = directory + salmon directory + sample name + quantifictaion file name

names(files) <- "~/Google Drive/Projects/color_patch_trimmed/gene counts/"
all(file.exists(files)) # do these all actually exist?
list(files)

#### make sample spreadsheet####

samples <- as.data.frame(filenames)

# get sample
samples$sample <- filenames %>% gsub(pattern = "_", replacement = "") %>% gsub(pattern = ".gene.counts", replacement = "") %>% gsub(pattern = "Black", replacement = "") %>% gsub(pattern = "Yellow", replacement = "") %>% gsub(pattern = "Varadero", replacement = "") %>% gsub(pattern = "Sauce", replacement = "")  %>% gsub(pattern = "Red", replacement = "")
samples$sample

# get tussue
samples$tissue <- filenames %>% gsub(pattern = "_", replacement = "") %>% gsub(pattern = ".gene.counts", replacement = "") %>% gsub(pattern = "[0-9]", replacement = "") %>% gsub(pattern = "Sauce", replacement = "") %>% gsub(pattern = "Varadero", replacement = "") %>% gsub(pattern = "Yellow", replacement = "Orange")
samples$tissue

# get morph
samples$morph <- filenames %>% gsub(pattern = "_", replacement = "") %>% gsub(pattern = ".gene.counts", replacement = "") %>% gsub(pattern = "Black", replacement = "") %>% gsub(pattern = "Yellow", replacement = "") %>% gsub(pattern = "[0-9]", replacement = "") %>% gsub(pattern = "Red", replacement = "") 
samples$morph

# save as spreadsheet
dir.create("~/Google Drive/Projects/color_patch_trimmed/results")
#write.csv(samples, "~/Google Drive/Color patch/results/samplespreadsheet.csv", row.names = F)

deseqsamples <- samples [,-1]

list(deseqsamples)

write.csv(deseqsamples, "~/Google Drive/Projects/color_patch_trimmed/results/samplespreadsheet.csv", row.names = F)
```

# Import expression data from.
```{r}
annos <- fread("Ranitomeya_imitator.imitator.1.3.6.annotations.genesymbol.tsv", header = FALSE)
colnames(annos) <- c("transcript", "Gene")
annos$Gene <- tolower(annos$Gene)

# rname columns from annos file
# import data
countdata <- data.frame(annos$transcript)
colnames(countdata) <- "transcript"
for (counts in (1:length(filenames))) {
  # import data for that sample
  toimport <- filenames[counts]
tmpdata <- read.table(paste0("gene counts/", toimport), header = FALSE, sep = "\t")
  # sample id
  samplename <- toimport %>% gsub(pattern = ".gene.counts", replacement = "")
  colnames(tmpdata) <- c("transcript", samplename)
  countdata <- dplyr::left_join(countdata, tmpdata, by = "transcript")
}


dim(countdata)
countdata <- data.frame(countdata, row.names = 1)
head(countdata)
head(annos)
```

# Import a priori candidate color gene
```{r}
colors <- read.csv("new_color_genes.csv", header = TRUE)
colnames(colors) <- "gene_name" 
colors$gene_name <- tolower(colors$gene_name)
head(colors)
```

# Quick data control. Change and "NA" values to "0". Also remove any row with total expression <50 (or 1 count/every 3rd sample).
```{r}
# change all NA to 0
print(paste("Number of NA after import:", sum(is.na(countdata))))
countdata[is.na(countdata)] <-  0
print(paste("Number of NA after NA removal:", sum(is.na(countdata))))

# remove any row with total expression < 10
print(paste("Number of rows (transcripts) after import:", nrow(countdata)))
keep <- rowSums(countdata) >= 50
countdata <- countdata[keep,]
print(paste("Number of rows (transcripts) after filter:", nrow(countdata)))

head(countdata)
countdata
write.csv(countdata, "~/Google Drive/Projects/color_patch_trimmed/results/count_data.csv", row.names = F)
```

```{r}
# make rownames column 1
countdata <- setDT(countdata, keep.rownames = "transcript")[]

# add gene symbol
countdata2 <- dplyr::left_join(annos, countdata, by = "transcript")

# change all NA to 0
print(paste("Number of NA after import:", sum(is.na(countdata2))))
countdata2[is.na(countdata2)] <-  0
print(paste("Number of NA after NA removal:", sum(is.na(countdata2))))


# combine all counts that map to the same gene
countdata3 <- aggregate(countdata2[, 3:18], list(countdata2$Gene), sum)
countdata3[is.na(countdata3)] <-  0
colnames(countdata3)[1] <- "Gene"
countdata <- countdata3

countdata <- data.frame(countdata, row.names = 1)

#newcountdata<- countdata[, c("Sauce_Black_15", "Sauce_Black_29")]

# save gene-level count data
write.table(countdata, "results/gene.level.count.data.tsv", row.names = TRUE, sep = "\t")
fdsafdsafdasgf
```

```{r}
dds <- DESeqDataSetFromMatrix (countData = countdata,
                                       colData = deseqsamples,
                                       design = ~ tissue + morph)
vsd <- vst(dds)



```

# create a PCA
```{r}
# get PC1 and PC2 data

pcaData <- plotPCA(vsd, intgroup = c("tissue", "morph"), returnData = TRUE)
pcaData
summary(pcaData)

plotPCA(vsd, intgroup=c("morph", "tissue"))

# get percent variation
percentVar <- round(100 * attr(pcaData, "percentVar"))

# pca code
ggplot(pcaData, aes(x = PC1, y = PC2, color = tissue, shape = morph ))+
  stat_ellipse(aes(group = morph), type = "t", level = .95, size = .55, show.legend = FALSE) + geom_point(size = 3.5, show.legend = TRUE) + 
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  coord_fixed() + theme_bw() + scale_shape_manual(values = c(18:15)) + scale_color_manual(values=c ("BLACK", "ORANGE", "RED")) + labs(color = "skin ccolor patch")  +  labs(shape = "color morph")  +  ggsave("results/figures/pca.png") 

#pcaExplorer(dds=dds)
```

#Only varadero morph 
```{r}
varadero.skincount <- dplyr::select(countdata, contains("Varadero"))
varadero.skinmorph <- deseqsamples %>% filter(morph == "Varadero")


# create DESeq dataset
dds.varadero.skin <- DESeqDataSetFromMatrix (countData = varadero.skincount, 
                                       colData = varadero.skinmorph,
                                       design = ~ tissue)

as.data.frame( colData(dds.varadero.skin) )

#run Wald test
dds.varadero.redVSblack <- DESeq(dds.varadero.skin, parallel=TRUE, BPPARAM=SnowParam(8))
res.varadero.redVSblack <- results(dds.varadero.redVSblack, parallel=TRUE, BPPARAM=SnowParam(8))
res.varadero.redVSblack$transcript <- mcols(res.varadero.redVSblack)$transcript

res.varadero.redVSblack

# how many are "significant"?
table(res.varadero.redVSblack[,"padj"] < 0.05)

#color genes 
SigGeneWrapper(res.varadero.redVSblack, 0.05, "res.varadero.redVSblack")
```


--------------
```{r}

varadero.skincount <- dplyr::select(countdata, contains("Varadero"))
varadero.skinmorph <- deseqsamples %>% filter(morph == "Varadero")


malanin_varadero.significant.color.dds <- DESeqDataSetFromMatrix (countData = varadero.skincount,
                                       colData = varadero.skinmorph, 
                                       design = ~ tissue)


#Melanin

varadero.melanin<- c("dct", "gpnmb", "gpr143", "kcnj13", "mc1r", "mitf", "oca2", "pax3", "pax7", "pmel", "rab27a", "slc24a5", "slc45a2", "tyrp1")


melanin_varadero.significant.color.dds <- DESeqDataSetFromMatrix (countData = varadero.skincount,
                                       colData = varadero.skinmorph, 
                                       design = ~ tissue)

melanin_varadero.significant.color.vsd <- varianceStabilizingTransformation(melanin_varadero.significant.color.dds, blind=FALSE)

melanin_varadero.significant.color.mat<- assay(melanin_varadero.significant.color.vsd)

melanin_varadero.significant.color.dds <- DESeq(melanin_varadero.significant.color.dds)

melanin_varadero.significant.color.res <- results(melanin_varadero.significant.color.dds)

melanin_varadero.significant.color.mat<- assay(melanin_varadero.significant.color.vsd)
#carotenoid_sauce.significant.color.mat <- carotenoid_sauce.significant.color.mat - rowMeans(carotenoid_sauce.significant.color.mat)
melanin_varadero.significant.color.df <- as.data.frame(colData(melanin_varadero.significant.color.vsd)[,c("tissue","morph")])

pheatmap(melanin_varadero.significant.color.mat[varadero.melanin,], cluster_cols = F, cluster_rows = T, scale = "row",labels_col = c("","Varadero: Black","","","varadero: Red",""),main = "Melanin/melanosome synthesis", border_color="Black", color = colorRampPalette(c("navy", "white", "firebrick3"))(50), angle_col = "0",annotation_legend = F, gaps_col = c(4),fontsize= 15, filename = "results/figures/varadero/publication/melanin_varadero_blackvsorange.png")


##################

varadero.iridophore<- c("gch1", "xdh", "pnp", "pts" )

iridophore_varadero.significant.color.dds <- DESeqDataSetFromMatrix (countData = varadero.skincount,
                                       colData = varadero.skinmorph, 
                                       design = ~ tissue)

iridophore_varadero.significant.color.vsd <- varianceStabilizingTransformation(iridophore_varadero.significant.color.dds, blind=FALSE)

iridophore_varadero.significant.color.mat<- assay(iridophore_varadero.significant.color.vsd)

iridophore_varadero.significant.color.dds <- DESeq(iridophore_varadero.significant.color.dds)

iridophore_varadero.significant.color.res <- results(iridophore_varadero.significant.color.dds)

iridophore_varadero.significant.color.mat<- assay(iridophore_varadero.significant.color.vsd)
#carotenoid_sauce.significant.color.mat <- carotenoid_sauce.significant.color.mat - rowMeans(carotenoid_sauce.significant.color.mat)
iridophore_varadero.significant.color.df <- as.data.frame(colData(iridophore_varadero.significant.color.vsd)[,c("tissue","morph")])

pheatmap(iridophore_varadero.significant.color.mat[varadero.iridophore,], cluster_cols = F, cluster_rows = T, scale = "row",labels_col = c("","Varadero: Black","","","varadero: Red",""),main = "Iridophore and pteridine", border_color="Black", color = colorRampPalette(c("navy", "white", "firebrick3"))(50), angle_col = "0",annotation_legend = F, gaps_col = c(4),fontsize= 15, filename = "results/figures/varadero/publication/iridophoresandguanine_varadero_blackvsorange.png")

  
  
###########
varadero.carotenoid<- c("dgat2", "mab21l2", "rbp1", "rbp2", "ttc39b")


carotenoid_varadero.significant.color.dds <- DESeqDataSetFromMatrix (countData = varadero.skincount,
                                       colData = varadero.skinmorph, 
                                       design = ~ tissue)

carotenoid_varadero.significant.color.vsd <- varianceStabilizingTransformation(carotenoid_varadero.significant.color.dds, blind=FALSE)

carotenoid_varadero.significant.color.mat<- assay(carotenoid_varadero.significant.color.vsd)

carotenoid_varadero.significant.color.dds <- DESeq(carotenoid_varadero.significant.color.dds)

carotenoid_varadero.significant.color.res <- results(carotenoid_varadero.significant.color.dds)


#carotenoid_sauce.significant.color.mat <- carotenoid_sauce.significant.color.mat - rowMeans(carotenoid_sauce.significant.color.mat)
carotenoid_varadero.significant.color.df <- as.data.frame(colData(carotenoid_varadero.significant.color.vsd)[,c("tissue","morph")])

pheatmap(carotenoid_varadero.significant.color.mat[varadero.carotenoid,], cluster_cols = F, cluster_rows = T, scale = "row",labels_col = c("","Varadero: Black","","","varadero: Red",""),main = "Carotenoid metabolism", border_color="Black", color = colorRampPalette(c("navy", "white", "firebrick3"))(50), angle_col = "0",annotation_legend = F, gaps_col = c(4),fontsize= 15, filename = "results/figures/varadero/publication/carotenoid_varadero_blackvsorange.png")

```

-----------------
#Only sauce morph
-----------------


```{r}
sauce.skincount <- dplyr::select(countdata, contains("Sauce"))
sauce.skinmorph <- deseqsamples %>% filter(morph == "Sauce")


# create DESeq dataset
dds.sauce.skin <- DESeqDataSetFromMatrix (countData = sauce.skincount, 
                                       colData = sauce.skinmorph,
                                       design = ~ tissue)

as.data.frame( colData(dds.sauce.skin) )

#run Wald test
dds.sauce.yellowVSblack <- DESeq(dds.sauce.skin, parallel=TRUE, BPPARAM=SnowParam(8))
res.sauce.yellowVSblack <- results(dds.sauce.yellowVSblack, parallel=TRUE, BPPARAM=SnowParam(8))
res.sauce.yellowVSblack$transcript <- mcols(res.sauce.yellowVSblack)$transcript

res.sauce.yellowVSblack

# how many are "significant"?
table(res.sauce.yellowVSblack[,"padj"] < 0.05)

#color genes 
SigGeneWrapper(res.sauce.yellowVSblack, 0.05, "ores.sauce.yellowVSblack")
```


```{r}

sauce.melanin<- c("mc1r","tyrp1","slc45a2","slc24a5","rab27a","pmel","pax7","pax3","oca2","mreg","mlph","mlana","kcnj13","gpr143","gpnmb", "erbb3", "dct", "trpm1")

sauce.skincount <- dplyr::select(countdata, contains("Sauce"))
sauce.skinmorph <- deseqsamples %>% filter(morph == "Sauce")

melanin_sauce.significant.color.dds <- DESeqDataSetFromMatrix (countData = sauce.skincount,
                                       colData = sauce.skinmorph, 
                                       design = ~ tissue)

melanin_sauce.significant.color.vsd <- varianceStabilizingTransformation(melanin_sauce.significant.color.dds, blind=FALSE)

melanin_sauce.significant.color.mat<- assay(melanin_sauce.significant.color.vsd)

melanin_sauce.significant.color.dds <- DESeq(melanin_sauce.significant.color.dds)

melanin_sauce.significant.color.res <- results(melanin_sauce.significant.color.dds)

melanin_sauce.significant.color.df <- as.data.frame(colData(melanin_sauce.significant.color.vsd)[,c("tissue","morph")])


pheatmap(melanin_sauce.significant.color.mat[sauce.melanin,], cluster_cols = F, cluster_rows = T, scale = "row",labels_col = c("","Sauce: Black","","","Sauce: Orange",""),main = "Melanin/melanosome synthesis", border_color="Black", color = colorRampPalette(c("navy", "white", "firebrick3"))(50), angle_col = "0",annotation_legend = F, gaps_col = c(4),fontsize= 15, filename = "results/figures/sauce/publication/melanin_sauce_blackvsorange.png")



##############

sauce.iridophore<- c("adsl","ak1", "atic", "gas1","gmps","paics","pnp","ppat", "pts", "xdh")


iridophore_sauce.significant.color.dds <- DESeqDataSetFromMatrix (countData = sauce.skincount,
                                       colData = sauce.skinmorph, 
                                       design = ~ tissue)


iridophore_sauce.significant.color.vsd <- varianceStabilizingTransformation(iridophore_sauce.significant.color.dds, blind=FALSE)

iridophore_sauce.significant.color.mat<- assay(iridophore_sauce.significant.color.vsd)

iridophore_sauce.significant.color.dds <- DESeq(iridophore_sauce.significant.color.dds)

iridophore_sauce.significant.color.res <- results(iridophore_sauce.significant.color.dds)

iridophore_sauce.significant.color.df <- as.data.frame(colData(iridophore_sauce.significant.color.vsd)[,c("tissue","morph")])


pheatmap(iridophore_sauce.significant.color.mat[sauce.iridophore,], cluster_cols = F, cluster_rows = T, scale = "row",labels_col = c("","Sauce: Black","","","Sauce: Orange",""),main = "Iridophore and pteridine", border_color="Black", color = colorRampPalette(c("navy", "white", "firebrick3"))(50), angle_col = "0",annotation_legend = F, gaps_col = c(4),fontsize= 15, filename = "results/figures/sauce/publication/iridophoresandguanine_sauce_blackvsorange.png")

#############

sauce.carotenoid<- c("dgat2","mab21l2", "rbp1", "scarb2", "ttc39b")


carotenoid_sauce.significant.color.dds <- DESeqDataSetFromMatrix (countData = sauce.skincount,
                                       colData = sauce.skinmorph, 
                                       design = ~ tissue)


carotenoid_sauce.significant.color.vsd <- varianceStabilizingTransformation(carotenoid_sauce.significant.color.dds, blind=FALSE)

carotenoid_sauce.significant.color.mat<- assay(carotenoid_sauce.significant.color.vsd)

carotenoid_sauce.significant.color.dds <- DESeq(carotenoid_sauce.significant.color.dds)

carotenoid_sauce.significant.color.res <- results(carotenoid_sauce.significant.color.dds)

carotenoid_sauce.significant.color.df <- as.data.frame(colData(carotenoid_sauce.significant.color.vsd)[,c("tissue","morph")])


pheatmap(carotenoid_sauce.significant.color.mat[sauce.carotenoid,], cluster_cols = F, cluster_rows = T, scale = "row",labels_col = c("","Sauce: Black","","","Sauce: Orange",""),main = "carotenoid", border_color="Black", color = colorRampPalette(c("navy", "white", "firebrick3"))(50), angle_col = "0",annotation_legend = F, gaps_col = c(4),fontsize= 15, filename = "results/figures/sauce/publication/carotenoid_sauce_blackvsorange.png")

```


display_numbers = TRUE
__________________________________






