---
title: "color patch trimmed orange vs red"
author: "Andrew O. Rubio"
date: "9/12/2020"
output: html_document
---

```{r}
library(tximport)
library(DESeq2)
library(dplyr)
library(foreach)
library(data.table)
library(splines)
library(ggthemes)
library(scales)
library(gridExtra)
library(tidyr)
library(pheatmap)
library(RColorBrewer)
library(ggplot2)
library("BiocParallel")
register(SnowParam(8))
library(apeglm)
library(tidyverse)
library(topGO)
library(GO.db)
library(tidyverse)


```


Some functions for this analysis.
```{r}

# this function pulls out particular a priori genes of interest into a new data frame
aprioris <- function(df, annos){
  df.aprioris <- data.frame()
  for (i in 1:length(colors$gene_name)){
    searchterm <- paste0("\\b", colors$gene_name[i], "\\b")
    searchterm <- tolower(searchterm)
    tmp <- df %>% filter(str_detect(Gene, searchterm))
    df.aprioris <- rbind(df.aprioris, tmp)
}
   return(df.aprioris)
}

# this function extracts all genes below a specified alpha value, prints the number of DE genes to screen, and saves a spreadsheet to file
siggies <- function(df, alpha, csv){
  sigs <- df %>% filter(padj < alpha)
  print(paste0("Number of significant genes: ", length(sigs$padj)))
  write.csv(sigs, csv, row.names = FALSE)
  return(sigs)
}


# significant gene wrapper function
SigGeneWrapper <- function(model_output, alpha, comparison){
  print(paste0("Running ", comparison, " with alpha = ", alpha))
  df <- setDT(as.data.frame(model_output), keep.rownames = "Gene")
  # add annotation data
  #df1 <- dplyr::left_join(df, annos, by = "transcript") 
  # get significant genes
  print("Overall significant genes")
  sigs <- siggies(df, alpha, paste0("results/results/", comparison, "_genes.csv"))
  # add annotation data
  colordf <- aprioris(df, annos)
  #get significant color genes
  print("Significant color genes")
  color.sigs <- siggies(colordf, alpha, paste0("results/results/", comparison, "_colorgenes.csv"))
}

```
#Data import

```{r}
sessionInfo()
```

#Data import

create sample spreadsheet.
```{r}
###### list all samples from expression data ####
# get the directory/path for each sample in this study
base_dir <- getwd()

filenames <- list.files(path = "color gene counts/", full.names = F, recursive = F)

files <- file.path(base_dir, "color gene counts/", filenames) # files = directory + salmon directory + sample name + quantifictaion file name

names(files) <- "~/Google Drive/Projects/color_patch_trimmed/color gene counts/"
all(file.exists(files)) # do these all actually exist?
list(files)

#### make sample spreadsheet####

color_samples <- as.data.frame(filenames)

# get sample
color_samples$sample <- filenames %>% gsub(pattern = "_", replacement = "") %>% gsub(pattern = ".gene.counts", replacement = "") %>% gsub(pattern = "Black", replacement = "") %>% gsub(pattern = "Yellow", replacement = "") %>% gsub(pattern = "Varadero", replacement = "") %>% gsub(pattern = "Sauce", replacement = "")  %>% gsub(pattern = "Orange", replacement = "")

# get tussue
color_samples$tissue <- filenames %>% gsub(pattern = "_", replacement = "") %>% gsub(pattern = ".gene.counts", replacement = "") %>% gsub(pattern = "[0-9]", replacement = "") %>% gsub(pattern = "Sauce", replacement = "") %>% gsub(pattern = "Varadero", replacement = "")
color_samples$tissue

# get morph
color_samples$morph <- filenames %>% gsub(pattern = "_", replacement = "") %>% gsub(pattern = ".gene.counts", replacement = "") %>% gsub(pattern = "Black", replacement = "") %>% gsub(pattern = "Yellow", replacement = "") %>% gsub(pattern = "[0-9]", replacement = "") %>% gsub(pattern = "Orange", replacement = "") 
color_samples$morph

# save as spreadsheet

color_deseqsamples <- color_samples [,-1]

list(color_deseqsamples)

write.table(color_deseqsamples, "~/Google Drive/Projects/color_patch_trimmed/Orange vs red/color_deseqsamples.tsv", row.names = TRUE, sep = "\t")

write.csv(color_deseqsamples, "~/Google Drive/Projects/color_patch_trimmed/results/color_samplespreadsheet.csv", row.names = F)
```

# Import expression data from.
```{r}
annos <- fread("Ranitomeya_imitator.imitator.1.3.6.annotations.genesymbol.tsv", header = FALSE)
colnames(annos) <- c("transcript", "Gene")
annos$Gene <- tolower(annos$Gene)

# rname columns from annos file
# import data
color_countdata <- data.frame(annos$transcript)
colnames(color_countdata) <- "transcript"
for (counts in (1:length(filenames))) {
  # import data for that sample
  toimport <- filenames[counts]
tmpdata <- read.table(paste0("color gene counts/", toimport), header = FALSE, sep = "\t")
  # sample id
  samplename <- toimport %>% gsub(pattern = ".gene.counts", replacement = "")
  colnames(tmpdata) <- c("transcript", samplename)
  color_countdata <- dplyr::left_join(color_countdata, tmpdata, by = "transcript")
}


dim(color_countdata)
color_countdata <- data.frame(color_countdata, row.names = 1)
head(color_countdata)
```

# Import a priori candidate color gene
```{r}
colors <- read.csv("new_color_genes.csv", header = TRUE)
colnames(colors) <- "gene_name" 
colors$gene_name <- tolower(colors$gene_name)
head(colors)
```

# Quick data control. Change and "NA" values to "0". Also remove any row with total expression <50 (or 1 count/every 3rd sample).
```{r}
# change all NA to 0
print(paste("Number of NA after import:", sum(is.na(color_countdata))))
color_countdata[is.na(color_countdata)] <-  0
print(paste("Number of NA after NA removal:", sum(is.na(color_countdata))))

# remove any row with total expression < 10
print(paste("Number of rows (transcripts) after import:", nrow(color_countdata)))
keep <- rowSums(color_countdata) >= 50
color_countdata <- color_countdata[keep,]
print(paste("Number of rows (transcripts) after filter:", nrow(color_countdata)))

head(color_countdata)
color_countdata
write.csv(color_countdata, "~/Google Drive/Projects/color_patch_trimmed/results/color_count_data.csv", row.names = F)

```

```{r}
# make rownames column 1
color_countdata <- setDT(color_countdata, keep.rownames = "transcript")[]

# add gene symbol
color_countdata2 <- dplyr::left_join(annos, color_countdata, by = "transcript")

# change all NA to 0
print(paste("Number of NA after import:", sum(is.na(color_countdata2))))
color_countdata2[is.na(color_countdata2)] <-  0
print(paste("Number of NA after NA removal:", sum(is.na(color_countdata2))))


# combine all counts that map to the same gene
color_countdata3 <- aggregate(color_countdata2[, 3:10], list(color_countdata2$Gene), sum)
color_countdata3[is.na(color_countdata3)] <-  0
colnames(color_countdata3)[1] <- "Gene"
color_countdata <- color_countdata3

color_countdata <- data.frame(color_countdata, row.names = 1)

# save gene-level count data
write.table(color_countdata, "results/color_gene.level.count.data.tsv", row.names = TRUE, sep = "\t")
fdsaffdaf

```



```{r}
dds_orangeVSred <- DESeqDataSetFromMatrix (countData = color_countdata,
                                       colData = color_deseqsamples,
                                       design = ~ tissue)
vsd_orangeVSred <- vst(dds_orangeVSred)

```

# create a PCA
```{r}

# get PC1 and PC2 data
pcaData <- plotPCA(vsd_orangeVSred, intgroup = c("tissue", "morph"), returnData = TRUE)
pcaData

plotPCA(vsd_orangeVSred, intgroup=c("morph", "tissue"))

# get percent variation
percentVar <- round(100 * attr(pcaData, "percentVar"))

# pca code
ggplot(pcaData, aes(x = PC1, y = PC2, color = tissue, shape = morph ))+
  stat_ellipse(aes(group = tissue), type = "t", level = 1., size = .75, show.legend = FALSE) + geom_point(size = 4, show.legend = TRUE) + 
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  coord_fixed() + theme_bw() + scale_shape_manual(values = c(18:15)) + scale_color_manual(values=c ("BLACK", "RED", "ORANGE")) + labs(color = "skin color patch")  +  labs(shape = "color morph")  +  ggsave("results/figures/color_pca.pdf") 

```

```{r}
#run Wald test
dds.color.orangeVSred <- DESeq(dds_orangeVSred, parallel=TRUE, BPPARAM=SnowParam(8))
res.color.orangeVSred <- results(dds.color.orangeVSred, parallel=TRUE, BPPARAM=SnowParam(8))
res.color.orangeVSred$transcript <- mcols(res.color.orangeVSred)$transcript

# how many are "significant"?
table(res.color.orangeVSred[,"padj"] < 0.05)

#color genes 
SigGeneWrapper(res.color.orangeVSred, 0.05, "res.color.orangeVSred")
```

```{r}
color.melanin<- c("adcy7", "bbs1", "bbs5", "creb3l1", "dct", "egfr", "erbb3", "gfpt1", "herc2", "lef1", "mitf", "mlana", "mlph", "mreg", "pmel", "rabggta", "sfxn1", "tyr", "tyrp1", "vps18", "wnt10b", "wnt7a")

melanin_color.significant.color.dds <- DESeqDataSetFromMatrix (countData = color_countdata ,
                                       colData = color_deseqsamples, 
                                       design = ~ tissue)


melanin_color.significant.color.vsd <- varianceStabilizingTransformation(melanin_color.significant.color.dds, blind=FALSE)

melanin_color.significant.color.mat<- assay(melanin_color.significant.color.vsd)

melanin_color.significant.color.dds <- DESeq(melanin_color.significant.color.dds)

melanin_color.significant.color.res <- results(melanin_color.significant.color.dds)

melanin_color.significant.color.df <- as.data.frame(colData(melanin_color.significant.color.vsd)[,c("tissue","morph")])

pheatmap(melanin_color.significant.color.mat[color.melanin,], cluster_cols = F, cluster_rows = T, scale = "row",labels_col = c("","Sauce: Orange","","","varadero: Red",""),main = "Melanin/melanosome synthesis", border_color="Black", color = colorRampPalette(c("navy", "white", "firebrick3"))(50), angle_col = "0",annotation_legend = F, gaps_col = c(4),fontsize= 15, filename = "results/figures/orangevsred/publication/melanin_ovsr_blackvsorange.png")

###############


color.iridophore<- c("adsl", "ak1", "anxa1", "dock7", "gart", "gchfr", "impdh2", "paics", "pgm2", "recql4", "rpia", "shmt2")


iridophore_color.significant.color.dds <- DESeqDataSetFromMatrix (countData = color_countdata ,
                                       colData = color_deseqsamples, 
                                       design = ~ tissue)


iridophore_color.significant.color.vsd <- varianceStabilizingTransformation(iridophore_color.significant.color.dds, blind=FALSE)

iridophore_color.significant.color.mat<- assay(iridophore_color.significant.color.vsd)

iridophore_color.significant.color.dds <- DESeq(iridophore_color.significant.color.dds)

iridophore_color.significant.color.res <- results(iridophore_color.significant.color.dds)

iridophore_color.significant.color.df <- as.data.frame(colData(iridophore_color.significant.color.vsd)[,c("tissue","morph")])

pheatmap(iridophore_color.significant.color.mat[color.iridophore,], cluster_cols = F, cluster_rows = T, scale = "row",labels_col = c("","Sauce: Orange","","","varadero: Red",""),main = "Iridophore and pteridine", border_color="Black", color = colorRampPalette(c("navy", "white", "firebrick3"))(50), angle_col = "0",annotation_legend = F, gaps_col = c(4),fontsize= 15, filename = "results/figures/orangevsred/publication/carotenoid_ovsr_blackvsorange.png")


########

  
color.carotenoid<- c("aldh1a1","crabp2", "mab21l2", "rbp1", "rbp2", "ttc8")


carotenoid_color.significant.color.dds <- DESeqDataSetFromMatrix (countData = color_countdata,
                                       colData = color_deseqsamples, 
                                       design = ~ tissue)

  
carotenoid_color.significant.color.vsd <- varianceStabilizingTransformation(carotenoid_color.significant.color.dds, blind=FALSE)

carotenoid_color.significant.color.mat<- assay(carotenoid_color.significant.color.vsd)

carotenoid_color.significant.color.dds <- DESeq(carotenoid_color.significant.color.dds)

carotenoid_color.significant.color.res <- results(carotenoid_color.significant.color.dds)

carotenoid_color.significant.color.df <- as.data.frame(colData(carotenoid_color.significant.color.vsd)[,c("tissue","morph")])

pheatmap(carotenoid_color.significant.color.mat[color.carotenoid,], cluster_cols = F, cluster_rows = T, scale = "row",labels_col = c("","Sauce: Orange","","","varadero: Red",""),main = "Carotenoid Metabolism", border_color="Black", color = colorRampPalette(c("navy", "white", "firebrick3"))(50), angle_col = "0",annotation_legend = F, gaps_col = c(4),fontsize= 15, filename = "results/figures/orangevsred/publication/carotenoid_ovsr_blackvsorange.png")

```



--------

```{r}
orangeVSred.significant.color.dds <- DESeqDataSetFromMatrix (countData = color_countdata,
                                       colData = color_deseqsamples, 
                                       design = ~ tissue)

as.data.frame( colData(orangeVSred.significant.color.dds) )




#run Wald test
dds.color.orangeVSred <- DESeq(orangeVSred.significant.color.dds, parallel=TRUE, BPPARAM=SnowParam(8))
res.color.orangeVSred <- results(dds.color.orangeVSred, parallel=TRUE, BPPARAM=SnowParam(8))
res.color.orangeVSred$transcript <- mcols(res.color.orangeVSred)$transcript

# how many are "significant"?
table(res.color.orangeVSred[,"padj"] < 0.05)

#color genes 
SigGeneWrapper(res.color.orangeVSred, 0.05, "res.color.orangeVSred")
```

```{r}
orangeVSred.significant.color.vsd <- varianceStabilizingTransformation(orangeVSred.significant.color.dds, blind=FALSE)

orangeVSred.significant.color.dds <- DESeq(orangeVSred.significant.color.dds)


# get PC1 and PC2 data
pcaData <- plotPCA(orangeVSred.significant.color.vsd, intgroup = c("tissue", "morph"), returnData = TRUE)
pcaData

plotPCA(orangeVSred.significant.color.vsd, intgroup=c("morph", "tissue"))

# get percent variation
percentVar <- round(100 * attr(pcaData, "percentVar"))


# pca code
ggplot(pcaData, aes(x = PC1, y = PC2, color = tissue, shape = morph ))+
  stat_ellipse(aes(group = tissue), type = "t", level = 0.95, size = .75, show.legend = FALSE) + geom_point(size = 1.5, show.legend = TRUE) + 
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  coord_fixed() + theme_bw() 
```

_____
