---
title: "color patch black vs color"
author: "Andrew O. Rubio"
date: "10/22/2020"
output: html_document
---

---
title: "color patch trimmed black vs color"
author: "Andrew O. Rubio"
date: "10/20/2020"
output: html_document
---


```{r}
library(tximport)
library(DESeq2)
library(dplyr)
library(foreach)
library(data.table)
library(splines)
library(ggthemes)
library(scales)
library(gridExtra)
library(tidyr)
library(pheatmap)
library(RColorBrewer)
library(ggplot2)
library("BiocParallel")
register(SnowParam(8))
library(apeglm)
library(tidyverse)
library(topGO)
library(GO.db)
library(tidyverse)
library(SummarizedExperiment)
library(pcaExplorer)
```


Some functions for this analysis.

```{r}
# this function pulls out particular a priori genes of interest into a new data frame
aprioris <- function(df, annos){
  df.aprioris <- data.frame()
  for (i in 1:length(colors$gene_name)){
    searchterm <- paste0("\\b", colors$gene_name[i], "\\b")
    searchterm <- tolower(searchterm)
    tmp <- df %>% filter(str_detect(Gene, searchterm))
    df.aprioris <- rbind(df.aprioris, tmp)
}
   return(df.aprioris)
}

# this function extracts all genes below a specified alpha value, prints the number of DE genes to screen, and saves a spreadsheet to file
siggies <- function(df, alpha, csv){
  sigs <- df %>% filter(padj < alpha)
  print(paste0("Number of significant genes: ", length(sigs$padj)))
  write.csv(sigs, csv, row.names = FALSE)
  return(sigs)
}


# significant gene wrapper function
SigGeneWrapper <- function(model_output, alpha, comparison){
  print(paste0("Running ", comparison, " with alpha = ", alpha))
  df <- setDT(as.data.frame(model_output), keep.rownames = "Gene")
  # add annotation data
  #df1 <- dplyr::left_join(df, annos, by = "transcript") 
  # get significant genes
  print("Overall significant genes")
  sigs <- siggies(df, alpha, paste0("results/results/", comparison, "_genes.csv"))
  # add annotation data
  colordf <- aprioris(df, annos)
  #get significant color genes
  print("Significant color genes")
  color.sigs <- siggies(colordf, alpha, paste0("results/results/", comparison, "_colorgenes.csv"))
}

```

#Data import

```{r}
sessionInfo()
```




```{r}
###### list all samples from expression data ####
# get the directory/path for each sample in this study
base_dir <- getwd()

filenames <- list.files(path = "bvsc gene counts/", full.names = F, recursive = F)

files <- file.path(base_dir, "bvsc gene counts/", filenames) # files = directory + salmon directory + sample name + quantifictaion file name

names(files) <- "~/Google Drive/Projects/color_patch_trimmed/bvsc gene counts/"
all(file.exists(files)) # do these all actually exist?
list(files)

#### make sample spreadsheet for black and color####

samples_color <- as.data.frame(filenames)

# get sample
samples_color$sample <- filenames %>% gsub(pattern = "_", replacement = "") %>% gsub(pattern = ".gene.counts", replacement = "") %>% gsub(pattern = "Black", replacement = "") %>% gsub(pattern = "One", replacement = "") %>% gsub(pattern = "Varadero", replacement = "") %>% gsub(pattern = "Sauce", replacement = "") %>% gsub(pattern = "Red", replacement = "") %>% gsub(pattern = "Yellow", replacement = "")

# get morph
samples_color$morph <- filenames %>% gsub(pattern = "_", replacement = "") %>% gsub(pattern = ".gene.counts", replacement = "") %>% gsub(pattern = "Black", replacement = "") %>% gsub(pattern = "Yellow", replacement = "") %>% gsub(pattern = "[0-9]", replacement = "") %>% gsub(pattern = "Red", replacement = "") %>% gsub(pattern = "One", replacement = "")
samples_color$morph

# get tussue
samples_color$tissue <- filenames %>% gsub(pattern = "_", replacement = "") %>% gsub(pattern = ".gene.counts", replacement = "") %>% gsub(pattern = "[0-9]", replacement = "") %>% gsub(pattern = "Sauce", replacement = "") %>% gsub(pattern = "Varadero", replacement = "") %>% gsub(pattern = "Yellow", replacement = "Color") %>% gsub(pattern = "Red", replacement = "Color") %>% gsub(pattern = "One", replacement = "")
samples_color$tissue

deseqsamples_color <- samples_color [,-1]

list(deseqsamples_color)

write.table(deseqsamples_color, "~/Google Drive/Projects/color_patch_trimmed/black vs color/allblackVScolordeseqsamples.tsv", row.names = TRUE, sep = "\t")

```





# Import expression data from.
```{r}
annos <- fread("Ranitomeya_imitator.imitator.1.3.6.annotations.genesymbol.tsv", header = FALSE)
colnames(annos) <- c("transcript", "Gene")
annos$Gene <- tolower(annos$Gene)

# rname columns from annos file
# import data
countdata <- data.frame(annos$transcript)
colnames(countdata) <- "transcript"
for (counts in (1:length(filenames))) {
  # import data for that sample
  toimport <- filenames[counts]
tmpdata <- read.table(paste0("bvsc gene counts/", toimport), header = FALSE, sep = "\t")
  # sample id
  samplename <- toimport %>% gsub(pattern = ".gene.counts", replacement = "")
  colnames(tmpdata) <- c("transcript", samplename)
  countdata <- dplyr::left_join(countdata, tmpdata, by = "transcript")
}


dim(countdata)
countdata <- data.frame(countdata, row.names = 1)
head(countdata)
head(annos)
```

# Import a priori candidate color gene
```{r}
colors <- read.csv("new_color_genes.csv", header = TRUE)
colnames(colors) <- "gene_name" 
colors$gene_name <- tolower(colors$gene_name)
head(colors)
```

# Quick data control. Change and "NA" values to "0". Also remove any row with total expression <50 (or 1 count/every 3rd sample).
```{r}
# change all NA to 0
print(paste("Number of NA after import:", sum(is.na(countdata))))
countdata[is.na(countdata)] <-  0
print(paste("Number of NA after NA removal:", sum(is.na(countdata))))

# remove any row with total expression < 10
print(paste("Number of rows (transcripts) after import:", nrow(countdata)))
keep <- rowSums(countdata) >= 50
countdata <- countdata[keep,]
print(paste("Number of rows (transcripts) after filter:", nrow(countdata)))

head(countdata)
countdata
write.csv(countdata, "~/Google Drive/Projects/color_patch_trimmed/results/blackvscolor_count_data.csv", row.names = F)
```

```{r}
# make rownames column 1
countdata <- setDT(countdata, keep.rownames = "transcript")[]

# add gene symbol
countdata2 <- dplyr::left_join(annos, countdata, by = "transcript")

# change all NA to 0
print(paste("Number of NA after import:", sum(is.na(countdata2))))
countdata2[is.na(countdata2)] <-  0
print(paste("Number of NA after NA removal:", sum(is.na(countdata2))))


# combine all counts that map to the same gene
countdata3 <- aggregate(countdata2[, 3:18], list(countdata2$Gene), sum)
countdata3[is.na(countdata3)] <-  0
colnames(countdata3)[1] <- "Gene"
countdata <- countdata3

countdata <- data.frame(countdata, row.names = 1)

#newcountdata<- countdata[, c("Sauce_Black_15", "Sauce_Black_29")]

# save gene-level count data
write.table(countdata, "results/bvsc.gene.level.count.data.tsv", row.names = TRUE, sep = "\t")
jfdsalhfjdklakh
```






```{r}
# create DESeq dataset for black vs color
dds.all.blackVScolor <- DESeqDataSetFromMatrix (countData = countdata,
                                       colData = deseqsamples_color,
                                       design = ~ tissue)

as.data.frame( colData(dds.all.blackVScolor) )

#run Wald test
dds.all.blackVScolor <- DESeq(dds.all.blackVScolor, test="Wald", parallel=TRUE, BPPARAM=SnowParam(8))
res.all.blackVScolor <- results(dds.all.blackVScolor, parallel=TRUE, BPPARAM=SnowParam(8))
res.all.blackVScolor$transcript <- mcols(res.all.blackVScolor)$transcript

res.all.blackVScolor

# how many are "significant"?
table(res.all.blackVScolor[,"padj"] < 0.05)

#color genes 
SigGeneWrapper(res.all.blackVScolor, 0.05, "res.allblackVScolor")

```

```{r}

vsd.all.blackVScolor <- vst(dds.all.blackVScolor)

# get PC1 and PC2 data
pcaData <- plotPCA(vsd.all.blackVScolor, intgroup = c("tissue"), returnData = TRUE)
pcaData

plotPCA(vsd.all.blackVScolor, intgroup=c("tissue"))

# get percent variation
percentVar <- round(100 * attr(pcaData, "percentVar"))

# pca code
ggplot(pcaData, aes(x = PC1, y = PC2, color = tissue))+
  stat_ellipse(aes(group = tissue), type = "t", level = 1., size = .75, show.legend = FALSE) + geom_point(size = 4, show.legend = TRUE) + 
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  coord_fixed() + theme_bw() + scale_shape_manual(values = c(18:15)) + scale_color_manual(values=c ("BLACK", "RED")) + labs(color = "skin color patch")  +  labs(shape = "color morph")  +  ggsave("results/figures/pca.pdf") 
```


```{r}
bvsc.significant.color.dds <- DESeqDataSetFromMatrix (countData = countdata,
                                       colData = deseqsamples_color, 
                                       design = ~ tissue)



bvsc.significant.color.vsd <- varianceStabilizingTransformation(bvsc.significant.color.dds, blind=FALSE)

bvsc.significant.color.mat<- assay(bvsc.significant.color.vsd)

bvsc.significant.color.dds <- DESeq(bvsc.significant.color.dds)

bvsc.significant.color.res <- results(bvsc.significant.color.dds)

bvsc.significant.color.df <- as.data.frame(colData(bvsc.significant.color.vsd)[,c("tissue","morph")])

#melanin
bvsc.melanin<- c("dct", "gpnmb", "gpr143", "kcnj13", "mc1r", "mitf", "mlana", "mreg", "oca2", "pax3", "pax7", "pmel", "rab27a", "slc24a5", "slc45a2", "trpm1", "tyrp1", "wnt16")

pheatmap(bvsc.significant.color.mat[bvsc.melanin,], cluster_cols = F, cluster_rows = T, scale = "row",labels_col = c("","","All Black","","","","Color","",""),main = "Melanin/melanosome synthesis", border_color="Black", color = colorRampPalette(c("navy", "white", "firebrick3"))(50), angle_col = "0",annotation_legend = F, gaps_col = c(8), fontsize= 15)#, filename = "results/figures/blackvscolor/publication/melanin_bvsc_blackvsred.png")

#iridophore

bvsc.iridophores<- c("fh", "gmps", "pnp", "ppat", "pts", "gch1", "xdh")

pheatmap(bvsc.significant.color.mat[bvsc.iridophores,], cluster_cols = F, cluster_rows = T, scale = "row",labels_col = c("","","All Black","","","","Color","",""),main = "iridophores and pteridine", border_color="Black", color = colorRampPalette(c("navy", "white", "firebrick3"))(50), angle_col = "0",annotation_legend = F, gaps_col = c(8), fontsize= 15)#, filename = "results/figures/blackvscolor/publication/iiridophore_bvsc_blackvsred.png")

#carotenoid
bvsc.carotenoid<- c("dgat2", "mab21l2", "ttc39b")

pheatmap(bvsc.significant.color.mat[bvsc.carotenoid,], cluster_cols = F, cluster_rows = T, scale = "row",labels_col = c("","","All Black","","","","Color","",""),main = "carotenoid metabolism", border_color="Black", color = colorRampPalette(c("navy", "white", "firebrick3"))(50), angle_col = "0",annotation_legend = F, gaps_col = c(8), fontsize= 15)#, filename = "results/figures/blackvscolor/publication/carotenoid_bvsc_blackvsred.png")
```
