---
title: "WGCNA"
author: "Andrew O. Rubio"
date: "4/10/2023"
output: html_document
---

```{r global_options, include=FALSE}
# load packages required for analyses
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(tximport)
library(DESeq2)
library(foreach)
library(data.table)
library(splines)
library(ggthemes)
library(scales)
library(gridExtra)
library(pheatmap)
library(RColorBrewer)
library(ggplot2)
library(BiocParallel)
register(SnowParam(8)) # set the number of threads to something reasonable
library(apeglm)
library(topGO)
library(GO.db)
library(WGCNA)
library(flashClust)
library(nlme)
library(tximport)
library(DESeq2)
#library(dplyr)
library(foreach)
library(data.table)
library(splines)
library(ggthemes)
library(scales)
library(gridExtra)
#library(tidyr)
library(pheatmap)
library(RColorBrewer)
library(ggplot2)
library("BiocParallel")
register(SnowParam(8))
library(apeglm)
#library(tidyverse)
library(topGO)
library(GO.db)
#library(tidyverse)
```

#### WGCNA analyses
```{r}
WGCNAdata <- read.table("r-output/gene.level.count.data.tsv", header = TRUE, sep = "\t")
WGCNAsamples <- read.csv("WGCNA/WGCNA_blackvscolor.csv")
```


```{r}
dds <- DESeqDataSetFromMatrix(countData = WGCNAdata,
                                       colData = WGCNAsamples,
                                       design = ~ Black_vs_Color + Morph)
as.data.frame(colData(dds))
```


```{r}
## using counts imported via tximports from above
vsd <- vst(dds, blind = FALSE)
vsd <- assay(vsd)
# transpose, as required by WGCNA
vsd <- as.data.frame(t(vsd))
```

checking for outliers... 
```{r}
# check if there are gene outliers
gsg = goodSamplesGenes(vsd, verbose = 3)
gsg$allOK
```

# there are outliers ... must remove them
```{r}
if (!gsg$allOK)
   {if (sum(!gsg$goodGenes)>0)
       printFlush(paste("Removing genes:", paste(names(vsd)[!gsg$goodGenes], collapse= ", ")));
       if (sum(!gsg$goodSamples)>0)
           printFlush(paste("Removing samples:", paste(rownames(vsd)[!gsg$goodSamples], collapse=", ")))
       vsd= vsd[gsg$goodSamples, gsg$goodGenes]
}

gsg = goodSamplesGenes(vsd, verbose = 3)
gsg$allOK

```

Do they match up?
```{r}
# verify data and sample information align
table(WGCNAsamples$sample==rownames(vsd))


# make sample id row name...
WGCNAsamples <- data.frame(WGCNAsamples[,-1], row.names=WGCNAsamples[,1])
WGCNAsamples <- data.frame(WGCNAsamples[,-1], row.names=WGCNAsamples[,1])
```

cluster samples by expression
```{r}
A = adjacency(t(vsd),type="signed") # this calculates the whole network connectivity
k = as.numeric(apply(A,2,sum))-1 # standardized connectivity
Z.k = scale(k)
thresholdZ.k = -2.5 # often -2.5
outlierColor = ifelse(Z.k<thresholdZ.k,"red","black")
sampleTree = flashClust(as.dist(1-A), method = "average")
# Convert traits to a color representation where red indicates high values
# needs to be numeric

WGCNAsamples$Black_vs_Color <- as.numeric(WGCNAsamples$Black_vs_Color)
WGCNAsamples$Morph <- as.numeric(WGCNAsamples$Morph)

traitColors=data.frame(numbers2colors(WGCNAsamples,signed=FALSE))
dimnames(traitColors)[[2]]=paste(names(WGCNAsamples))
datColors=data.frame(outlier=outlierColor,traitColors)

plotDendroAndColors(sampleTree, groupLabels = names(datColors), colors = datColors, main="sample Dendrogram and Trait Heatmap")
```

soft threshold
```{r}
powers = c(c(1:10), seq(from =12, to=40, by=2)) #choosing a set of soft-thresholding powers
sft = pickSoftThreshold(vsd, powerVector=powers, verbose =5, networkType="signed") #call network topology analysis function
```


scale independence
```{r}
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], xlab= "Soft Threshold (power)", ylab="Scale Free Topology Model Fit, signed R^2", type= "n", main= paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], labels=powers, col="red")
abline(h=0.9, col="red")

plot(sft$fitIndices[,1], sft$fitIndices[,5], xlab= "Soft Threshold (power)", ylab="Mean Connectivity", type="n", main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, col="red")

```

################ Construct network ################

```{r}
net = blockwiseModules(vsd, power = 20,
    TOMType = "unsigned", minModuleSize = 30,
    reassignThreshold = 0, mergeCutHeight = 0.25,
    numericLabels = TRUE, pamRespectsDendro = FALSE,
    saveTOMs = TRUE,
    saveTOMFileBase = "MimeticGenesTOM",
    verbose = 3)
```

plot these modules
```{r}
table(net$colors)
# Convert labels to colors for plotting
mergedColors = labels2colors(net$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
    "Module colors",
    dendroLabels = FALSE, hang = 0.03,
    addGuide = TRUE, guideHang = 0.05)
```


```{r}
nGenes=ncol(vsd);
nSamples=nrow(vsd);

MEs0=moduleEigengenes(vsd, mergedColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, WGCNAsamples, use = "p");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);

textMatrix= paste(signif(moduleTraitCor, 2), "\n(", 
                    signif(moduleTraitPvalue, 1), ")", sep= "");
dim(textMatrix)= dim(moduleTraitCor)
par(mar= c(6, 8.5, 3, 3))

#display the corelation values with a heatmap plot
labeledHeatmap(Matrix= moduleTraitCor, 
            xLabels= names(WGCNAsamples), 
            yLabels= gsub(pattern = "ME",
replacement = "", names(MEs)), 
            ySymbols= names(MEs), 
            colorLabels= FALSE, 
            colors= blueWhiteRed(50), 
            textMatrix= textMatrix, 
            setStdMargins= FALSE, 
            cex.text= 1, 
            zlim= c(-1,1), 
            main= paste("Module-trait relationships"))

# save to disk

png("WGCNA/results/Figures/power20.blackvscolor.WGCNA_heatmap_new_putative_fix.png", width = 7, height = 12, units = "in", res = 300)
par(mar= c(6, 12, 3, 3))
labeledHeatmap(Matrix= moduleTraitCor, 
            xLabels= names(WGCNAsamples), 
            yLabels=  gsub(pattern = "ME",
replacement = "", names(MEs)), 
            ySymbols= names(MEs), 
            colorLabels= FALSE, 
            colors= blueWhiteRed(50), 
            textMatrix= textMatrix, 
            setStdMargins= FALSE, 
            cex.text= 1, 
            zlim= c(-1,1), 
            main= paste("Module-trait relationships"))

dev.off()
```

```{r WGCNA: output module membership}
datME=moduleEigengenes(vsd,mergedColors)$eigengenes
datKME=signedKME(vsd, datME, outputColumnName="MM.")
genes=names(vsd)
geneInfo0 = data.frame(gene=genes,moduleColor=mergedColors)
color=data.frame(geneInfo0,datKME) #these are from your original WGCNA analysis 
head(color)
write.csv(as.data.frame(color), file = "WGCNA/results/GeneModule_membership.csv")

#### MM pvalues
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(vsd, MEs, use = "p"));
MMPvalue=as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
pvals=data.frame(geneModuleMembership,MMPvalue)
head(pvals)
write.csv(as.data.frame(pvals), file = "WGCNA/results/GeneModule_membership_pvalues.csv")

```


PULL OUT datKME by module of interest, rank them high membership to lowest (from 1 to -1), then run through GO.

For each signifcant module I want to see:
1. What genes are in these modules.
2. If there are candidate color genes.
3. GO of this grouping.

### Significant gene modules
```{R, WGNCA: identify significant modules byx color patch}
# subets out only significant modules
modulePs <- as.data.frame(moduleTraitPvalue)
Black_vs_Colorsigmods <- rownames(subset(modulePs, Black_vs_Color < 0.05))
print(paste0("Number of significant modules by black_vs_color: ", length(Black_vs_Colorsigmods)))
morphsigmods <- rownames(subset(modulePs, Morph < 0.05))
print(paste0("Number of significant modules by morph: ", length(morphsigmods)))

```


#color patch
```{R Hub genes}
# subets out only significant modules
modulePs <- as.data.frame(moduleTraitPvalue)
Black_vs_Colorsigmods <- rownames(subset(modulePs, Black_vs_Color < 0.05))
print(paste0("Number of significant modules by color morph: ", length(Black_vs_Colorsigmods)))
tissuesigmods <- rownames(subset(modulePs, Black_vs_Color < 0.05))
print(paste0("Number of significant modules by color patch: ", length(Black_vs_Colorsigmods)))

sigmods <- rownames(modulePs[rowSums(modulePs < 0.05) >= 1, ])
print(paste0("Number of unique significant modules by any variable: ", length(sigmods)))


# remove "ME" from all sigmods
sigmods <- gsub(pattern = "ME", replacement = "", sigmods)
probes = names(vsd)
nTop = 7
moduleColors=mergedColors
adj = adjacency(vsd, power = 20, type = "signed") #specify network type

## Translate adjacency into topological overlap matrix and calculate the corresponding dissimilarity

TOM = TOMsimilarity(adj, TOMType="signed") #Topological Overlap Matrix
dissTOM = 1-TOM 

for(i in 1:length(unique(sigmods))){
 #inModule = sigmods[i];
 #modProbes = probes[inModule];
 inModule = (moduleColors == sigmods[i]);
 modProbes = probes[inModule];
 ## Select the corresponding Topological Overlap

 modTOM = TOM[inModule, inModule];
 dimnames(modTOM) = list(modProbes, modProbes)

 ## Genes with highest connectivity within each significant module

 IMConn = softConnectivity(vsd[, modProbes])
 top = (rank(-IMConn) <= nTop)
IMConn
 top
 
 tempdat<-data.frame(IMConn,modProbes)[order(-IMConn),]
 names(tempdat)<-c("IMConnectivity","target_id")
 tempdat$ModuleColor <- rep(sigmods[i],nrow(tempdat))
 #tempdat2 = left_join(tempdat,preGO1, by = "target_id")
 write.csv(tempdat, paste("WGCNA/results/hub/blackvscolor.HubGenes",sigmods[i],".csv",sep="_"), na="", row.names=FALSE)
}

## List all the files

temp <- list.files(path = "WGCNA/results/hub/", pattern = "*HubGenes")
temp <- paste("WGCNA/results/hub/", temp, sep = "")
Hub_genes = do.call(rbind, lapply(temp, function(x) read.csv(x, stringsAsFactors = FALSE)))
write.csv(Hub_genes, paste("WGCNA/results/hub/blackvscolor.HubGenes_all.csv",sep="_"), na="", row.names=FALSE)

```


```{R HUBGENES2}
# Recalculate topological overlap if needed


# Select modules
sigmods # Lists modules of interest
module = "brown"    


moduleColors = mergedColors
# Select module probes
probes = names(vsd)
inModule = is.finite(match(moduleColors, module));
modProbes = probes[inModule];
#modGenes = annote_df$gene_name[match(modProbes, annote_df$target_id)];

# Select the corresponding Topological Overlap
modTOM = TOM[inModule, inModule];
dimnames(modTOM) = list(modProbes, modProbes)
cyt = exportNetworkToCytoscape(modTOM,
      edgeFile = paste("WGCNA/results/CytoscapeInput-edges_colorvsblack-", paste(module, collapse="-"), ".txt"),
      nodeFile = paste("results/results/CytoscapeInput-nodes_colorvsblacl-", paste(module, collapse="-"), ".txt"),
      weighted = TRUE,
      threshold = .2807, # Change to get top 50 genes
      nodeNames = modProbes,
      #altNodeNames = modGenes,
      nodeAttr = moduleColors[inModule]);
cyt



cytoscapePing()

```

#Morph
```{R Hub genes}

# subets out only significant modules
modulePs <- as.data.frame(moduleTraitPvalue)
morphsigmods <- rownames(subset(modulePs, Morph < 0.05))
print(paste0("Number of significant modules by color morph: ", length(morphsigmods)))
tissuesigmods <- rownames(subset(modulePs, Black_vs_Color < 0.05))
print(paste0("Number of significant modules by color patch: ", length(morphsigmods)))

morphsigmods <- rownames(modulePs[rowSums(modulePs < 0.05) >= 1, ])
print(paste0("Number of unique significant modules by any variable: ", length(morphsigmods)))


# remove "ME" from all sigmods
morphsigmods <- gsub(pattern = "ME", replacement = "", morphsigmods)
probes = names(vsd)
nTop = 7
moduleColors=mergedColors
adj = adjacency(vsd, power = 20, type = "signed") #specify network type

## Translate adjacency into topological overlap matrix and calculate the corresponding dissimilarity

TOM = TOMsimilarity(adj, TOMType="signed") #Topological Overlap Matrix
dissTOM = 1-TOM 

for(i in 1:length(unique(morphsigmods))){
 #inModule = sigmods[i];
 #modProbes = probes[inModule];
 inModule = (moduleColors == morphsigmods[i]);
 modProbes = probes[inModule];
 ## Select the corresponding Topological Overlap

 modTOM = TOM[inModule, inModule];
 dimnames(modTOM) = list(modProbes, modProbes)

 ## Genes with highest connectivity within each significant module

 IMConn = softConnectivity(vsd[, modProbes])
 top = (rank(-IMConn) <= nTop)
IMConn
 top
 
 tempdat<-data.frame(IMConn,modProbes)[order(-IMConn),]
 names(tempdat)<-c("IMConnectivity","target_id")
 tempdat$ModuleColor <- rep(morphsigmods[i],nrow(tempdat))
 #tempdat2 = left_join(tempdat,preGO1, by = "target_id")
 write.csv(tempdat, paste("WGCNA/results/hub/blackvscolor.HubGenes",morphsigmods[i],".csv",sep="_"), na="", row.names=FALSE)
}

## List all the files

temp <- list.files(path = "WGCNA/results/hub/", pattern = "*HubGenes")
temp <- paste("WGCNA/results/hub/", temp, sep = "")
Hub_genes = do.call(rbind, lapply(temp, function(x) read.csv(x, stringsAsFactors = FALSE)))
write.csv(Hub_genes, paste("WGCNA/results/hub/blackvscolor.HubGenes_all.csv",sep="_"), na="", row.names=FALSE)

```


```{R HUBGENES2}

# Recalculate topological overlap if needed


# Select modules
morphsigmods # Lists modules of interest
module = "greenyellow"    


moduleColors = mergedColors
# Select module probes
probes = names(vsd)
inModule = is.finite(match(moduleColors, module));
modProbes = probes[inModule];
#modGenes = annote_df$gene_name[match(modProbes, annote_df$target_id)];

# Select the corresponding Topological Overlap
modTOM = TOM[inModule, inModule];
dimnames(modTOM) = list(modProbes, modProbes)
cyt = exportNetworkToCytoscape(modTOM,
      edgeFile = paste("WGCNA/results/CytoscapeInput-edges_morph-", paste(module, collapse="-"), ".txt"),
      nodeFile = paste("results/results/CytoscapeInput-nodes_morph-", paste(module, collapse="-"), ".txt"),
      weighted = TRUE,
      threshold = .157, # Change to get top 50 genes
      nodeNames = modProbes,
      #altNodeNames = modGenes,
      nodeAttr = moduleColors[inModule]);
cyt



cytoscapePing()

```